### 2023-06-07 Conference Submission/DraftDoc.tex
eoeu

\begin{section}{Introduction}
    Temporal networks are networks whose nodes and edges change in time. In this thesis we look to model temporal networks. That is, describe the temporal network using mathematical language. 
    % Why temp nets are important
    The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.

    % Temp nets are dynam systs
    Temporal networks can be seen as a system in which points, in our case nodes in a network, whose states, the edges connecting them, that vary dependent in time, a dynamical system.

    % Hard to find dynam systs because of discrete events 
    Discovering the underlying equations governing these dynamical systems proves challenging. That is because changes in network structure are typically observed in the form of discrete jumps from one state to another, for example an edge between two nodes not being observed at the first time step then being observed at the next.

    % We are proposing a solution
    In this thesis, we propose a hybrid statistical and deep learning framework. That is a framework in which a neural network is used to model probabilistic values of the temporal network. This framework allows us to model temporal networks as continuous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a network.

    % DEs are useful for modelling interacting states
    Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables.

    % Temp nets have interacting states 
    We observe this behavior in temporal networks; nodes' connections within the network can influence the observation of edges between other nodes, for example the phenomenon observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more likely to gain the connections the more connections it has (preferential attachment). With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

    % Temp net events are discrete
    In the common representation of networks as binary-valued adjacency matrices, matrices where the observation, or lack thereof, of an edge is represented as a 1 or 0 respectively, the events recorded in a temporal sequence of networks correspond to the appearance or the disappearance of link.

    % Difficult to model discrete events
    Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require us to handle discrete jumps.
    The topological nature of temporal networks, and the discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

    % We propose a solution
    Here, we overcome the discreteness problem by interpreting networks as a well established statistical model for complex networks, that embeds nodes in a continuous, low-dimensional metric space by using a truncated singular value decomposition (Random Dot Product Graphs\cite{athreya2017statistical}). In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continuous change in the embedding space.

    % Our solution is interpretable because we use symb reg
    We then define and use systems of Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to predict forward in time the evolution of the temporal networks.

    % We propose a solution
    In this manuscript, we show that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem by taking the singular value decomposition of a sequence of adjacency matrices and training a NNDE to model an approximation to the underlying differential equation. We then go on to create a symbolic equation of this approximation.

@@

    % The framework can be applied to lots of fields
    As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations.
\end{section}

@@

    \subsection{Temporal Networks}
        % Use of temp nets
        Temporal networks are used to describe a series of interactions across time.
        A temporal network $G$ is defined as:  
        \begin{align}
            G=\{ {_t} G=({_t} N, {_t} E)\} _{t \in 1\ldots T } 
        \end{align}
        $_t x \in {_t} N, {_t} e\in {_t} E$, where ${_t} x, {_t} e$ are nodes and edges at time $t$ respectively, and ${_t} N, {_t} E$ are the sets of nodes and edges at time $t$ respectively.
        An example of this can be found in \autoref{method nets}.

        %representations of temp nets as adj lists
        This can be represented in many ways, for example as a series of interactions. The series of interactions is generally represented as a list of tuples containing the node that interacted and the time that the interaction took place. 
        \begin{align}
            G=\{(i, j, t)|\space i,j\in {_t} N, (i\leftrightarrow j)\in {_t}E \}
        \end{align}

@@

        % Uses of temp nets
        Temporal networks are present in many areas of research interest, ranging from symptom interactions in mental health\cite{jordan2020current,contreras2020temporal}, to epidemiology\cite{masuda2013predicting}, protein interactions\cite{lucas2021inferring,jin2009identifying}, and social networks\cite{moinet2015burstiness,hanneke2010discrete}.

@@

        An example of modelling the change in a node state could be the severity of a symptom at a given time, as in \Citeauthor{contreras2020temporal}\cite{contreras2020temporal}. Where the node $_t x$ represents the symptom and $m_1$ would represent the severity of that symptom.

        Other effort focus on modelling the changing structure of the network as in \Citeauthor{sanna2021link}\cite{sanna2021link}, which looks to model $_t G$ as a whole.

        % Intro contreras as attempt to model changes in nodes
        \Citeauthor{contreras2020temporal}\cite{contreras2020temporal} model the change in node state over time while the structure of the network remains constant. \Citeauthor{contreras2020temporal} look at the severity of symptoms over time, i.e. the state of the nodes, as a temporal network where the network structure remains constant.

        % Contreras methods
        They use multilevel vector autoregression, a method that finds correlations between the current state of a variable and the states of variables at previous time steps \cite{singer2003applied}, to model how the current symptoms of people with paranoia might predict their future symptoms. These models were then used to create three networks that linked symptoms.

@@

        % intro to paper as attempt to model node state
        In \cite{KARIMI20133476}, the \Citeauthor{KARIMI20133476} look to model the change in node state while the network structure changes.

@@

        % what state the paper attempts to model
        The authors use a model to predict when a node will change from state 0 to state 1, in practice this could represent whether a user of a social network believes a rumour, or an ecological patch being colonised.

        % Model limitation
        Notably, in their model the node never changes back; this is done to keep the model analytically tractable, but does limit its usefulness in some contexts. In this model an organism could not become locally extinct in an ecological patch for example.

@@

        % Limitation of model
        In this work I aim to present a model for predicting the state of a node in the future. Instead, \Citeauthor{KARIMI20133476} aim to present a tool for understanding the spread of ideas, viruses, etc. within a network.

        % Intro to pasino et al as modelling structure of temp net
        \Citeauthor{sanna2021link}\cite{sanna2021link} model the change in the structure of a temporal network.

@@

        % What is a latent space
        A latent space is a continuous space in which similar nodes (nodes with many of the same neighbours) are positioned near each other. In this space nodes are represented as points in a continuous, low dimensional space. Another feature of this latent space, is that the dot product of any two points in the space gives us the probability that the nodes related to those points are connected.

@@

    \subsection{UDEs and NNDEs}
        % Justification for DEs being useful
        In some temporal networks the state of a node might influence the evolution of its neighbours. For example, in social networks we often observe the power law distribution of edges, where the more neighbours a node has, the more likely they are to gain more neighbours\cite{zhao2012multi,garg2009evolution}. We also see that the state of symptoms (occurrence, severity, and distress), can influence the states of other symptoms in patients undergoing chemotherapy \cite{papachristou2019network,kalantari2022network}, and for mental health disorders \cite{contreras2020temporal}. As well as in ecological networks where the interactions of species are often represented as a network, and the populations of one species may influence the population growth of another \cite{elton2001animal,volterra1927variazioni}; that is, the state (in this case the population) of a species (node), may influence the movement of another species.

        % DEs have lots of research
        With the observation that nodes in real world temporal networks may influence others, differential equations or difference equations seem appropriate for modelling the evolution of nodes as there is a rich body of literature exploring these methods use in modelling interacting variables.

        % The data we have is discrete, which is a problem for DEs
        One of the issues that the differential equation method encounters is the prevalence of discrete jumps of edges forming and decaying. Because of this we may consider exploring the possibility of using difference equations.

        % time is continuous so good to have our model be cont too
        We note that the progression of the network is generally continuous even if the events happen as discrete jumps. Because of this, we may wish to preserve the continuous temporal progression in our model.

@@

        % passino has discrete data but uses a method to view it as cont
        Looking at the work of Passino et al.\cite{sanna2021link}, the authors use random dot produce graphs\cite{athreya2017statistical} to approximate a temporal network as a matrix of probabilities of an edge existing between two nodes. In contrast to edges, these probabilities can evolve continuously, and so we can use differential equations to generate probabilistic networks at any temporal resolution we require. That is, we will be able to model the changes in the probabilities of edges continuous and from this, generate a network at any time step.

        %  we can use the same method we can view data as dynam sys and use DEs
        Using the method in Passino et al.\cite{sanna2021link}, we can treat the sequence of embeddings as a dynamical system and then use differential equation modelling in a continuous space. That is, once we have found a latent space for each network. We know that each point in this space refers to a specific node in the network, and so, as the network changes, so too does the position of each of the points in the latent space. We then look to model this movement using differential equations.

        % Can use a new NN structure to make this framework very general
        Our aim for this framework to be as general as possible and note that there may not yet be enough domain knowledge to perform traditional differential equation modelling to the desired accuracy. For example in network ecology models can be useful, but not perfect XXX cite XXX. Hence, we might wish to build on these models rather than trying to create something entirely new. This can be achieved by using this partial domain knowledge using it to create a universal differential equation (UDE) \cite{SciML_C_Rak}.

@@

            % Overview of why UDEs are useful
            Developed by \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, Universal Iifferential Equation (UDEs) are a novel neural network architecture that aims to take the best of both traditional neural differential equation modelling and of flexible machine learning approaches to modelling.

@@

            The UDE will incorporate some component of $f(t,y(t))\approx \hat f(t,y(t))$, and will learn the difference with a Neural Network (NN) such that:
            \begin{align}
                \label{UDE}
                y(0) &= y_0 \\
                \frac{dy}{dt}(t)&=\hat f(t,y(t))+NN(t,y(t), \theta)
            \end{align}
            Where $\theta$ is a vector of the parameters of the NN.
            % Structure of UDE
            UDEs combine domain knowledge in the form of a differential equation, with a neural network. This hybrid machine learning model is then trained on the observed data.

            % Overview of why UDEs are useful
            Combining the two allows for most of the movement in the data to be captured by the differential equation, which contains the domain knowledge. In this way the neural network will only need to learn a theoretically simpler equation, given that a part of the system has already been captured.

            % Why this structure is useful
            This then allows the neural network to be smaller, require fewer data, and be trained much faster than traditional neural network models. Given the flexibility of neural networks, the use of UDEs is natural when modelling a process that is not entirely understood\cite{kidger2022neural}.

            % Software used for coding
            There is an ecosystem of packages specifically designed for performing machine learning with UDEs. The ecosystem is called SciML or scientific machine learning\cite{SciML_C_Rak}. Largely, the SciML ecosystem has been optimised for flexibility and efficiency with respect to models available and training performance. Given there does not seem to be any other ecosystem with this feature, it seemed like 
            an obvious choice.

        \subsubsection{Applications}
            Whilst there has been much interest in implementing UDEs and neural ODEs, much of this work has been focussed on physics informed neural network and physics informed ordinary differential equations (PINNs and PINODEs) \cite{karniadakis2021physics,GAO2021110079,krishnapriyan2021characterizing,roehrl2020modeling}, and on improving modelling of fluids \cite{mahmoudabadbozchelou2021data,nguyen2022physics}. Alongside this, and given UDEs have gained popularity in the last few years, there have been a number of studies exploring their usefulness in modeling the effect of restrictions due to COVID-19 on the virus' spread \cite{Dandekar2020.04.03.20052084}. Although there has been a large amount of research into the usefulness of these types of models, to the best of this author's knowledge they have never been applied to the problem of predicting temporal networks. This seems to be a natural fit for UDEs and neural ODEs, given we may know relatively little about the processes that govern the evolution of temporal networks, and any information we do have can be incorporated into the model.

    \subsection{Symbolic Regression}
        % What is symbolic regression
        Symbolic regression is a type of analysis that looks to find a relationship between input variables that describe the output variables. For symbolic regression, we define a set of equations that may be used to find this relationship. The algorithm then generates combinations of these equations to find an expression that matches the output variables. Because we define the set of potential equations, the resulting expression is interpretable by humans, as opposed to a NN for example.

@@

        % How is this done
        This can be achieved in many ways and remains active area of research. One of the most popular algorithms is genetic programming\cite{schmidt2009distilling}.

        % How genetic programming works
        Genetic programming involves constructing a search tree using a genetic algorithm to yield a potential equation, which may match the data as well as numerically calculated partial derivatives of each of the input variables.

@@

        % Examples of use of genetic programming
        For example, both \Citeauthor{schmidt2009distilling,bongard2007automated} test their methods' capacity for obtaining a symbolic equation with physical simulation data, various pendula specifically. A ``meaningful'' equation in this case would be one that adheres to the laws of physics.

        % Expalaination of schmidt
        Examples of genetic programming include discovering equations governing the movement of simple harmonic and chaotic double pendula from data\cite{schmidt2009distilling}. In this paper the authors present an evolutionary algorithm to generate multiple candidate functions that approximate the partial derivatives of each variable that are calculated directly from the data.

        % Intro to bongard
        Another approach to symbolic regression can be found in \cite{bongard2007automated}. This paper proposes a method to find a symbolic equation that approximates a learned numerical equation rather than directly from data.

        % How does this process work
        The candidate equation is then tested against the numerical equation in such a way that the difference between the two predictions are maximised. The algorithm will generate new initial conditions to find behaviour exhibited by the numerical equation, for example a fixed point, that is not exhibited by the functional equation. When a difference is found, the functional equation is adjusted to also exhibit the behaviour of the numerical equation.

        % Limitations of this process
        Especially when using a neural network model as the numerical equation, it seems these process may not be useful if the initial conditions are outside the scope of the training data.

        % Benefits of combining UDEs and symbolic regression
        In very recent work, Kidger\cite{kidger2022neural}, demonstrates that, with UDEs, two of the assumptions necessary for symbolic regression can be overcome. Those are the necessity for paired observations and derivatives, as well as the assumption that the function can be expressed as a shallow tree of symbolic operations. We are modelling a differential equation, and so will have access to the derivatives of our numerical model at any point, as neural networks are easily differentiable.

        % Usefulness of symbolic regression for nnde
        In \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, the authors demonstrate that using symbolic regression, as opposed to just an NNDE, can improve the accuracy of extrapolating. One aspect of this improvement is having accurate numerical derivative to compare to. To ensure this, it is important that the data is sampled very frequently. This seems integral to the accuracy of the found equation. 
\end{section}

@@

        % How we achieve this goal
        To achieve this, we reinterpret each graph as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then Train a Neural Network Differential Equation (NNDE) to approximate the rate of change of the points in this latent space. With this numerical equation we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a differential equation.

        \begin{figure}
            \centering
            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.25\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology1.tex}}
                    \label{method net, a}
                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.25\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology2.tex}}
                    \label{method net, b}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.25\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology3.tex}}
                    \label{method net, c}
                \end{subfigure}

@@

            \begin{tabular}{llll}

            \end{tabular}

@@

    \subsection{Comparable Methods and Related Work}
    Teneto Python

    packages("sna")  R
    packages("tsna")
    packages("ndtv")

    EvolvingGraphs Julia

    \subsection{Past and Ongoing Research}
    Wrote code that is general

    Made a structure (TemporalNetworkEmbedding)

>>>>

\begin{section}{Introduction}
    Temporal networks are networks whose nodes and edges change in time. In this thesis we look to model temporal networks. That is, describe the changing temporal network using mathematical language. 
    % Why temp nets are important
    The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.

    % Temp nets are dynam systs
    Temporal networks can be seen as a system in which points, in our case nodes in a network, whose states, the edges connecting them, vary dependent in time, a dynamical system.

    % Hard to find dynam systs because of discrete events 
    Discovering the underlying equations governing these dynamical systems proves challenging. That is because changes in network structure are typically observed in the form of discrete jumps from one state to another; for example an edge between two nodes not being observed at the first time step then being observed at the next.

    % We are proposing a solution
    In this thesis, we propose a hybrid statistical and deep learning framework. That is a framework in which a neural network is used to model probabilistic values of the temporal network. This framework allows us to model temporal networks as continuous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a temporal network.

    % DEs are useful for modelling interacting states
    Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behavior in temporal networks; nodes' connections within the network can influence the observation of edges between other nodes, for example the phenomenon observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more likely to gain the connections the more connections it has (preferential attachment). With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

    % Temp net events are discrete
    In the common representation of networks as binary-valued adjacency matrices, matrices where the observation or lack thereof, of an edge is represented as a 1 or 0 respectively, the events recorded in a temporal sequence of networks correspond to the appearance or the disappearance of link.

    % Difficult to model discrete events
    Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require us to handle discrete jumps. The discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

    % We propose a solution
    Here, we overcome the discreteness problem by interpreting temporal networks as a well established statistical model for complex networks, which embeds nodes in a continuous, low-dimensional metric space by using a truncated singular value decomposition (Random Dot Product Graphs\cite{athreya2017statistical}). In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continuous change in the embedding space.

    % Our solution is interpretable because we use symb reg
    We then define and use Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to predict forward in time the evolution of the temporal networks.

    % We propose a solution
    In this manuscript, we show that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem by taking the singular value decomposition of a sequence of adjacency matrices and training a NNDE to model an approximation to the underlying differential equation. We then go on to create a functional equation of this approximation.

@@

    % The framework can be applied to lots of fields
    As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations (UDEs).
\end{section}

@@

    \subsection{Temporal Networks}
        % Use of temp nets
        \subsubsection{Definitions}
        Temporal networks are used to describe a series of interactions across time.
        A temporal network $G$ is defined as:  
        \begin{align}
            G=\{ {_t} G=({_t} N, {_t} E)\} _{t \in 1\ldots T } 
        \end{align}
        $_t x \in {_t} N, {_t} e\in {_t} E$, where ${_t} x, {_t} e$ are nodes and edges at time $t$ respectively, and ${_t} N, {_t} E$ are the sets of nodes and edges at time $t$ respectively.
        An example of this can be found in \autoref{method nets}.

        %representations of temp nets as adj lists
        This can be represented in many ways, for example as a series of interactions. The series of interactions is generally represented as a list of tuples containing the nodes which interacted and the time that the interaction took place. 
        \begin{align}
            G=\{(i, j, t)|\space i,j\in {_t} N, (i\leftrightarrow j)\in {_t}E \}
        \end{align}

@@

        \subsubsection{Applications}
        Temporal networks are present in many areas of research interest, ranging from symptom interactions in mental health\cite{jordan2020current,contreras2020temporal}, to epidemiology\cite{masuda2013predicting}, protein interactions\cite{lucas2021inferring,jin2009identifying}, and social networks\cite{moinet2015burstiness,hanneke2010discrete}.

@@

        An example of modelling the change in a node state could be the severity of a symptom at a given time, as in \Citeauthor{contreras2020temporal}\cite{contreras2020temporal}. Where the node $_t x$ represents a given symptom and $m_1$ would represent the severity of that symptom.

        % Intro contreras as attempt to model changes in nodes
        \Citeauthor{contreras2020temporal}\cite{contreras2020temporal} model the change in node state over time while the structure of the network remains constant. The authors look at the severity of symptoms over time, i.e. the state of the nodes, where the nodes exist in a statically structured temporal network.

        % Contreras methods
        \Citeauthor{contreras2020temporal} use multilevel vector autoregression, a method that finds correlations between the current state of a variable and the states of variables at previous time steps \cite{singer2003applied}, to model how the current symptoms of people with paranoia might predict their future symptoms. These models were then used to create three networks that linked symptoms.

@@

        % intro to paper as attempt to model node state
        In \cite{KARIMI20133476}\cite{KARIMI20133476}, the \Citeauthor{KARIMI20133476} look to model the change in node state while the network structure changes.

@@

        % what state the paper attempts to model
        The authors use a model to predict when a node will change from state 0 to state 1, in practice this could represent whether a user of a social network believes a rumour, or whether an ecological patch has been colonised.

        % Model limitation
        Notably, in this model, the node never changes back; this is done to keep the model analytically tractable, but does limit its usefulness in some contexts. In this model an organism could not become locally extinct in an ecological patch for example.

@@

        % Limitation of model
        In this work we aim to present a model for predicting the state of a node in the future. Instead, \Citeauthor{KARIMI20133476} aim to present a tool for understanding the spread of ideas, viruses, etc. within a network.

        Other efforts focus on modelling the changing structure of the network as in \Citeauthor{sanna2021link}\cite{sanna2021link}, which looks to model $_t G$ as a whole.

@@

        % What is a latent space
        A latent space is a continuous space in which similar nodes (nodes with many of the same neighbours) are positioned near each other. In this space nodes are represented as points in a continuous, low dimensional space. Another feature of this latent space, is that the dot product of any two points in the space gives us the probability that the nodes related to those points are connected, a Random Dot Product Graph (RDPG)\cite{athreya2017statistical}.

@@

    \subsection{UDEs and NNDEs}
        % Justification for DEs being useful
        In some temporal networks the state of a node might influence the evolution of its neighbours. For example, in social networks we often observe that the more neighbours a node has, the more likely they are to gain more neighbours, a power law distribution of edges\cite{zhao2012multi,garg2009evolution}. We also see that the state of symptoms (occurrence, severity, and distress), can influence the states of other symptoms in patients undergoing chemotherapy \cite{papachristou2019network,kalantari2022network}, and for mental health disorders \cite{contreras2020temporal}. As well as in ecological networks where the interactions of species are often represented as a network, and the populations of one species may influence the population growth of another \cite{elton2001animal,volterra1927variazioni}; that is, the state (in this case the population) of a species (node), may influence the movement of another species.

        % DEs have lots of research
        With the observation that nodes in real world temporal networks may influence others, differential equations or difference equations seem appropriate for modelling the evolution of nodes; there is a rich body of literature exploring these methods use in modelling interacting variables.

        % The data we have is discrete, which is a problem for DEs
        One of the issues that the differential equation method encounters is the prevalence of discrete jumps of edges being observed and then not being observed. Because of this, we may consider exploring the possibility of using difference equations.

        % time is continuous so good to have our model be cont too
        We note that the progression of the network is generally continuous even if the events happen as discrete jumps. We may wish to preserve this continuous temporal progression in our model.

@@

        % passino has discrete data but uses a method to view it as cont
        Looking at the work of \Citeauthor{sanna2021link}\cite{sanna2021link}, the authors use RDPGs to approximate a temporal network as a matrix of probabilities of an edge existing between two nodes. In contrast to edges, these probabilities can evolve continuously, and so we can use differential equations to generate probabilistic networks at any temporal resolution we require. That is, we will be able to model the changes in the probabilities of edges continuously and from this, generate a network at any time step.

        %  we can use the same method we can view data as dynam sys and use DEs
        Using the method in \Citeauthor{sanna2021link}\cite{sanna2021link}, we can treat the sequence of embeddings as a dynamical system and then use differential equation modelling in a continuous space. Once we have found a latent space for each network, we know that each point in this space refers to a specific node in the network, and so, as the network changes, so too does the position of each of the points in the latent space. We then look to model this movement using differential equations.

        % Can use a new NN structure to make this framework very general
        Our aim for this framework to be as general as possible and note that there may not yet be enough domain knowledge to perform traditional differential equation modelling to the desired accuracy. For example in network ecology models can be useful, but not perfect. Hence, we might wish to build on these models rather than trying to create something entirely new. This can be achieved by using this partial domain knowledge using it to create a UDE\cite{SciML_C_Rak}.

@@

            % Overview of why UDEs are useful
            Developed by \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, UDEs are a novel neural network architecture that aims to take the best of both traditional neural differential equation modelling and of flexible machine learning approaches to modelling.

@@

            The UDE will incorporate some component of $f(t,y(t))\approx \hat f(t,y(t))$, and will learn the difference with a Neural Network (NN) such that:
            \begin{align}
                \label{UDE}
                y(0) &= y_0 \\
                \frac{dy}{dt}(t)&=\hat f(t,y(t))+NN(t,y(t), \theta)
            \end{align}
            Where $\theta$ is a vector of the parameters of the NN.
            % Structure of UDE
            UDEs combine domain knowledge in the form of a differential equation, with a NN. This hybrid machine learning model is then trained on the observed data.

            % Overview of why UDEs are useful
            Combining the two allows for most of the movement in the data to be captured by the differential equation, which contains the domain knowledge. In this way, the NN will only need to learn a theoretically simpler equation, given that a part of the system has already been captured.

            % Why this structure is useful
            This then allows the NN to be smaller, require fewer data, and be trained much faster than traditional NN models. Given the flexibility of NNs, the use of UDEs is natural when modelling a process that is not entirely understood\cite{kidger2022neural}.

            Note that we can construct a special case of a UDE, an NNDE when $\hat f(t,y(t))=0$. This structure can be used when do not have any domain knowledge we wish to give the model.

            % Software used for coding
            There is an ecosystem of packages specifically designed for performing machine learning with UDEs. The ecosystem is called Scientific Machine Learning (SciML)\cite{SciML_C_Rak}. Largely, the SciML ecosystem has been optimised for flexibility and efficiency with respect to models available and training performance. Given there does not seem to be any other ecosystem with this feature, it seemed like 
            an obvious choice.

        \subsubsection{Applications}
            Whilst there has been much interest in implementing UDEs and NNDEs, much of this work has been focussed on physics informed neural network and physics informed neural ordinary differential equations (PINNs and PINODEs) \cite{karniadakis2021physics,GAO2021110079,krishnapriyan2021characterizing,roehrl2020modeling}, and on improving modelling of fluids \cite{mahmoudabadbozchelou2021data,nguyen2022physics}. Alongside this, UDEs have gained popularity in the last few years, and there have been a number of studies exploring their usefulness in modeling the effect of restrictions due to COVID-19 on the virus' spread \cite{Dandekar2020.04.03.20052084}. Although there has been a large amount of research into the usefulness of these types of models, to the best of this author's knowledge they have never been applied to the problem of predicting temporal networks. This seems to be a natural fit for UDEs and NNDEs, given we may know relatively little about the processes that govern the evolution of temporal networks, and any information we do have can be incorporated into the model.

    \subsection{Symbolic Regression}
        % What is symbolic regression
        Symbolic regression is a type of analysis that looks to take input variables and find a relationship between them that describes the output variables. For symbolic regression, we construct a set of mathematical operations that may be used to find this relationship. The algorithm then generates combinations of these equations to find an expression that matches the output variables. Because we construct the set of potential operations, the resulting expression is interpretable by humans (a functional equation), as opposed to a NN for example (a numeric equation).

@@

        % How is this done
        Finding a suitable functional expression can be achieved in many ways and remains active area of research. One of the most popular algorithms is genetic programming\cite{schmidt2009distilling}.

        % How genetic programming works
        Genetic programming involves constructing a search tree using an evolutionary algorithm to yield candidate equations. An evolutionary algorithm is an iterative process. It begins by randomly generating a set of candidate equations with operations from the given set of operations:
        \begin{align}
            Q_1=\{f_1,\dots,f_n\}
        \end{align}
        Each candidate function $f_i$ is tested and given a fitness score based on how well it performed. The higher a candidates' fitness, the more likely it is to be selected for the next iteration.

        Suppose $f_1,f_2$ were selected for the next iteration. The evolutionary algorithm would then randomly generate new functions based on $f_1,f_2$ again with the given set of operations.
        \begin{align}
            Q_2=\{f_{1,1},\dots,f_{1,n}, f_{2,1},\dots,f_{2,n}\}
        \end{align}
        This process is repeated until a suitable functional equation is found.

        For the genetic programming algorithm, the fitness of a candidate function is determined by how closely it matches the data as well as numerically calculated partial derivatives of each of the input variables.

@@

        % Examples of use of genetic programming
        For example, both \Citeauthor{schmidt2009distilling,bongard2007automated} test their methods' capacity for obtaining a functional equation with physical simulation data, various pendula specifically. A ``meaningful'' equation in this case may be an equation that conserves momentum.

        % Expalaination of schmidt
        Examples of genetic programming include discovering equations governing the movement of simple harmonic and chaotic double pendula from data\cite{schmidt2009distilling}. In \Citeauthor{schmidt2009distilling}, the authors present an evolutionary algorithm to generate multiple candidate functions that approximate the partial derivatives of each variable that are calculated directly from the data.

        % Intro to bongard
        Another approach to symbolic regression can be found in \Citeauthor{bongard2007automated}\cite{bongard2007automated}. The authors propose a method to find a functional equation that approximates a learned numerical equation rather than directly from data.

        % How does this process work
        The algorithm generates a set of candidate equations. The candidate equations are then tested against the numerical equation in such a way that the difference between the candidate and numerical predictions are maximised. The algorithm will generate new initial conditions to find behaviour exhibited by the numerical equation, for example a fixed point, that is not exhibited by the functional equation. When a difference is found, the functional equations are adjusted to also exhibit the behaviour of the numerical equation.

        % Limitations of this process
        We will be using a UDE with relatively small data sets. The models will be trained on a small range of input values, and so we would not have confidence that any behaviour exhibited outside the range of these input values would reflect the underlying system. As such, this method may not yield us useful results.

        % Benefits of combining UDEs and symbolic regression
        In very recent work, Kidger\cite{kidger2022neural} demonstrates that, with UDEs, two of the assumptions necessary for symbolic regression can be overcome. Those are the requirement for paired observations and derivatives, as well as the assumption that the function can be expressed as a shallow tree of symbolic operations. We are modelling a differential equation, and so will have access to the derivatives of our numerical model at any point, as neural networks are easily differentiable.

        % Usefulness of symbolic regression for nnde
        In \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, the authors demonstrate that by using symbolic regression to find a functional equation of our model, as opposed to just a UDE, we can improve the accuracy of extrapolating.
\end{section}

@@

        % How we achieve this goal
        To achieve this, we reinterpret our temporal network as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then train a Neural Network Differential Equation (NNDE) to approximate the rate of change of the points in this latent space. With this numerical equation we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a temporal network.

        \begin{figure}
            \centering
            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology1.tex}}
                    \label{method net, a}
                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology2.tex}}
                    \label{method net, b}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology3.tex}}
                    \label{method net, c}
                \end{subfigure}

@@

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                    $\begin{bmatrix}
                        -0.82  & 0.01\\
                        -0.64  & 0.36\\
                        -0.95 & -0.76\\
                        -0.64 &  0.36\\
                        -0.48 &  0.82\\
                        -0.17 & -0.40\\
                        -0.20 & -0.52\\
                        -0.07 &  0.25\\
                    \end{bmatrix}$

                &

                    $\begin{bmatrix}
                        -0.82  &  0.10\\
                        -0.82  &  0.10\\
                        -0.90 & -0.58\\
                        -0.82  &  0.10\\
                        -0.42 &  0.81\\
                        -0.13  & -0.64\\
                        -0.13  & -0.64\\
                        -0.12 &  0.57\\
                    \end{bmatrix}$

                &
                $\cdots$
                &

                    $\begin{bmatrix}
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                         0.08 &  -0.69\\
                         0.11 &  -0.88\\
                         0.11 &  -0.88\\
                         0.08 &  -0.69\\
                    \end{bmatrix}$
                \end{tabular}

            \end{subfigure}

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 1.png}
                        \label{}
                    \end{subfigure}

                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 2.png}
                        \label{}
                    \end{subfigure}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 3.png}
                        \label{}
                    \end{subfigure}
                \end{subfigure}

            \end{subfigure}

@@

    \subsection{Future Work}
        The current TNE structure is not yet GPU compatible. We plan to remedy this as well as include support for constructing a TNE from Graph.jl and MetaGraph.jl objects.
